Classless Java: Tuning Java Interfaces

### 2016-03-10 OOPSLA16/sections/Formalization.tex

\section{Interaction of Interface Methods with Interface Composition}
\bruno{I don't think this section makes alot of sense in the context 
of the new paper!}
Before formalizing Classless Java and object interfaces, it is helpful
to informally discuss the behaviour of Java8
interfaces, in contrast with conventional trait models.

>>>>

\section{Interaction of Interface Methods with Interface Composition}
\bruno{I don't think this section makes alot of sense in the context 
of the new paper! I think perhaps some of this could be mentioned 
in related work.

In my opinion, instead of this section, we could have one section 
that demonstrates our implementation and its advantages.

\begin{itemize}

\item Show a diagram of how the approach works?

\item Syntax and type system reuse; (good for error messages, for example)
\item On-the-fly AST rewriting (means no new files are created, and functionality
is immediatelly available for IDE's for example); this is better than conventional 
annotation processors.
\item Eclipse integration; code completion.
\item No integration issues in typical pre-processors; (the process is transparent; 
no need for an external tool that generates code in a first step).
\item Separate compilation (which relates to the properties?)

\end{itemize}

}
Before formalizing Classless Java and object interfaces, it is helpful
to informally discuss the behaviour of Java8
interfaces, in contrast with conventional trait models.

### 2016-03-10 OOPSLA16/sections/Implementation.tex
OOPSLA format.

The annotation we created is \mixin. In Eclipse, with an interface annotated by
\mixin, the automatic annotation processing is performed transparently and the information of
the interface from compilation is captured in the ``Outline'' window. This includes
all the methods inside the interface as well as the generated ones.  The custom
transformation is easy and convenient to use.  For example this means that the
IDE functionality for content assist and autocomplete will work for the newly generated
methods. The biggest reasons to use Lombok rather than using a conventional Java
 annotation processor are:
\begin{itemize}
%this seems like a commercial spot
%\item Lombok is byte-code based instead of source-code based, which makes client
%  code concise and easy to maintain. Such code generation is performed at
%  compile time to modify bytecode.
\item Lombok modifies the generation process of the class files, by directly modifying the AST.
Neither the source code is modified nor new Java files are generated.
\item Moreover, and probably more importantly, Lombok is capable of generating
  code \emph{inside} a class/interface, which conventional
  Java annotation processors do not support.
\end{itemize}

>>>>

The annotation we created is \mixin. In Eclipse, with an interface annotated by
\mixin, the automatic annotation processing is performed transparently and the information of
the interface from compilation is captured in the ``Outline'' window. This includes
all the methods inside the interface as well as the generated ones.  The custom
transformation is easy and convenient to use.  For example, this means that the
IDE functionality for content assist and autocomplete will work for the newly generated
methods. The biggest reasons to use Lombok rather than using a conventional Java
 annotation processor are:
\begin{itemize}
%this seems like a commercial spot
%\item Lombok is byte-code based instead of source-code based, which makes client
%  code concise and easy to maintain. Such code generation is performed at
%  compile time to modify bytecode.
\item Lombok modifies the generation process of the class files, by directly modifying the AST.
Neither the source code is modified nor new Java files are generated.
\item Moreover, and probably more importantly, Lombok is capable of generating
  code \emph{inside} a class/interface, which conventional
  Java annotation processors do not support.
\end{itemize}

\noindent \textcolor{red}{Another big concern of using Lombok in our approach is that, the modularity, code quality and maintainability
are highly improved in contrast with preprocessors. For the last decades, we have seen a lot of convenience in
programs using preprocessors; it is easy to learn and use, lightweight in code size, and applicable in various
languages including C, Java, and so on. Nevertheless, preprocessors are heavily criticized in literature in
terms of code quality and maintainability being affected and subtle errors being introduced. In [ref], the main
aspects of criticism are discussed including:}

\begin{itemize}
\item \textbf{Separation of concerns.} \textcolor{red}{Preprocessors can affect the modularity of implementation code. Whereas in our approach,
each feature can be realized by a single annotation with its handler, and meanwhile a declared Java class or interface
can be applied with multiple annotations, which ensures the modularity.}
\item \textbf{Sensitivity to subtle errors.} \textcolor{red}{Preprocessors like cpp, which introduces the syntax of ``\#ifdef'', are likely to
bring very simple syntax errors or type errors. In our Lombok-based approach, annotation processing is applied between
parsing and type-checking, where internal ASTs are modified. Lombok uses an internal AST for Java, with respect to
Java syntax, hence syntax errors are avoided. On the other hand, possible type errors introduced by Lombok handlers
can be observed by Java type system from compilation, which brings convenience in practice.}
\item \textbf{Obfuscated source code.} \textcolor{red}{Implementations with preprocessors obfuscate the source code, making code less readable
and hard to maintain with further changes. In our case, since modularity is supported, the handler code can be debugged
and maintained for different features respectively. On the other hand, annotations do not introduce much confusion in
the source code, while the information of annotated interface from compilation can still be captured in the "Outline" window
of Eclipse or other IDEs.}
\item \textbf{Lack of reuse.} \textcolor{red}{Reusability is yet another concern in using preprocessors. In Lombok, implementations of features are
encapsulated in various annotation handlers, in which case some behaviours are allowed to reuse the code by invoking methods
in other handlers, where tedious replicated code is avoided.}
\end{itemize}
